Detecting and quantifying parasite-induced host mortality from intensity data: method comparisons and limitations

### 2015-06-12 docs/parasite_mortality_manuscript.tex
Moved figures out of main text and into separate file for journal formatting

% Format the Bibliography appropriately
% increase \bibhang to take care of the numbers
\setlength{\bibhang}{0pc}
\makeatletter
% patch \@lbibitem to print the current number before the authors
\patchcmd{\@lbibitem}
  {]}
  {] \theNAT@ctr. \newline }
  {}{}
\makeatother

@@

\title{Detecting and quantifying parasite-induced host mortality from intensity data: method comparisons and limitations}
\author{Mark Wilber, Sara Weinstein, and Cherie Briggs}

@@

\section{Methods}

@@

Generally, a larger $b$ leads to a more rapid decline in the probability of host survival as parasite intensity increases, with the maximum rate of decline having a value of $b / 4$ (\emph{SI} 2). $b$ is in many ways analogous to the the pathogenicity parameter ($\alpha$) in classic macroparasite models that gives the parasite intensity dependent host death rate. \citep{AndersonandMay1978,Isham1995}. When $b$ is held constant, a larger $a$ allows for hosts to tolerate larger parasite intensities before experiencing parasite-induced mortality. More specifically, for every one unit increase in $a$ the log parasite intensity at which any percent of hosts survive (e.g. 99\% of hosts survive) increases by $ 1 / b$ (\emph{SI} 2).

@@

We thank the Briggs Lab group and Theoretical Ecology group at UCSB for helpful feedback.  This work was supported by National Science Foundation Graduate Research Fellowship (Grant No. DGE 1144085) and X.

@@

\renewcommand{\arraystretch}{1.2}

\begin{table}
    \centering
    \caption{Definition of parameters and functions used in the main text}
    \begin{tabular}{c p{12cm}}
    \hline
    Parameter & Definition \\
    \hline\hline
    $\mu_p$ & Pre-mortality mean parasite intensity \\
    $k_p$   & Pre-mortality parasite aggregation \\
    $N_p$   & Pre-mortality host sample size \\
    $x$     & Number of parasites in a given host \\
    $g(x; \mu_p, k_p)$ & Pre-mortality negative binomial parasite distribution \\

    $h(\text{survival}; x, a, b)$ & The probability of host survival given a parasite load $x$ and logistic parameters $a$ and $b$\\

    $b / 4$ & The maximum rate of decline in host survival probability with increasing parasite intensity  \\
    $a$ & When $b$ is held constant a one unit increase in $a$ leads to a $1 / b$ increase in the log parasite intensity at which some percent of hosts survive\\
    $LD_{50}$ & $\exp(a / b)$, parasite intensity above which a host has a greater than 50\% chance of dying \\

    \end{tabular}
    \label{tab:params}
\end{table}

@@

    \begin{tikzpicture}
        \node at (0, 0) (hist) {\includegraphics[width=\textwidth]{/Users/mqwilber/Repos/parasite_mortality/results/concept_crofton_method}};

    \end{tikzpicture}

@@

\begin{tikzpicture}

    \node at (0, 0) (pic) {\includegraphics[width=\textwidth]{/Users/mqwilber/Repos/parasite_mortality/results/figure1_partII_for_manuscript50}};

    \node[above=0.1cm] at (pic.north) (concept) {\includegraphics[width=0.6\textwidth]{/Users/mqwilber/Repos/parasite_mortality/results/figure1_partI_for_manuscript}};

    \draw[->] (0, 4.2)--(-1.5, 3.2);
    \draw[->] (0, 4.2)--(1.5, 3.2);

    \node at (-3.2, 4.9) {A.};
    \node at (-6.7, 2.9) {B.};
    \node at (1.15, 2.9) {C.};

\end{tikzpicture}

@@

\begin{figure}
    \includegraphics[width=\textwidth]{/Users/mqwilber/Repos/parasite_mortality/results/figure2_for_manuscript.pdf}

@@

\begin{figure}
    \includegraphics[width=\textwidth]{/Users/mqwilber/Repos/parasite_mortality/results/figure3_for_manuscript.pdf}

>>>>

% Format the Bibliography appropriately
% increase \bibhang to take care of the numbers
% \setlength{\bibhang}{0pc}
% \makeatletter
% % patch \@lbibitem to print the current number before the authors
% \patchcmd{\@lbibitem}
%   {]}
%   {] \theNAT@ctr. \newline }
%   {}{}
% \makeatother

@@

@@

\section{Materials and methods}

@@

Generally, a larger $b$ leads to a more rapid decline in the probability of host survival as parasite intensity increases, with the maximum rate of decline having a value of $b / 4$ (\emph{SI} 2). $b$ is in many ways analogous to the the pathogenicity parameter ($\alpha$) in classic macroparasite models that gives the parasite intensity dependent host death rate. \citep{AndersonandMay1978,Isham1995}. When $b$ is held constant, a larger $a$ allows for hosts to tolerate larger parasite intensities before experiencing parasite-induced mortality. More specifically, for every one unit increase in $a$ the log parasite intensity at which any percent of hosts survive (e.g. 99\% of hosts survive) increases by $1 / b$ (\emph{SI} 2).

@@

We thank the Briggs Lab group and Theoretical Ecology group at UCSB for helpful feedback.  MW was supported by National Science Foundation Graduate Research Fellowship (Grant No. DGE 1144085) and the UC Regents CB was supported by X.

@@

% \renewcommand{\arraystretch}{1.2}

% \begin{table}
%     \centering
%     \caption{Definition of parameters and functions used in the main text}
%     \begin{tabular}{c p{12cm}}
%     \hline
%     Parameter & Definition \\
%     \hline\hline
%     $\mu_p$ & Pre-mortality mean parasite intensity \\
%     $k_p$   & Pre-mortality parasite aggregation \\
%     $N_p$   & Pre-mortality host sample size \\
%     $x$     & Number of parasites in a given host \\
%     $g(x; \mu_p, k_p)$ & Pre-mortality negative binomial parasite distribution \\

%     $h(\text{survival}; x, a, b)$ & The probability of host survival given a parasite load $x$ and logistic parameters $a$ and $b$\\

%     $b / 4$ & The maximum rate of decline in host survival probability with increasing parasite intensity  \\
%     $a$ & When $b$ is held constant a one unit increase in $a$ leads to a $1 / b$ increase in the log parasite intensity at which some percent of hosts survive\\
%     $LD_{50}$ & $\exp(a / b)$, parasite intensity above which a host has a greater than 50\% chance of dying \\

%     \end{tabular}
%     \label{tab:params}
% \end{table}

@@

    %\includegraphics[width=\textwidth]{/Users/mqwilber/Repos/parasite_mortality/results/concept_crofton_method}
    \captionsetup{justification=raggedright, singlelinecheck=false}

@@

    \captionsetup{justification=raggedright, singlelinecheck=false}

@@

    %\includegraphics[width=\textwidth]{/Users/mqwilber/Repos/parasite_mortality/results/figure2_for_manuscript.pdf}

    \captionsetup{justification=raggedright, singlelinecheck=false}

@@

\begin{figure}
    %\includegraphics[width=\textwidth]{/Users/mqwilber/Repos/parasite_mortality/results/figure3_for_manuscript.pdf}

    \captionsetup{justification=raggedright, singlelinecheck=false}
